---
title: "Capstone Project Data Analyst"
author: Lucas Bagge
date: '2022-11-01'
format: 
  html:
    code-fold: true
    theme: 
      - yeti
---

```{r setup}
#| include: false
library(knitr)
knitr::opts_chunk$set(cache = TRUE,
                      cache.lazy = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      echo = TRUE,
                      dpi = 180,
                      fig.width = 7,
                      fig.height = 5
                      #fig.asp = 0.88,
                      #out.width = "90%"
)
library(tidyverse)
library(ggplot2)
library(tidymodels)
library(brms)
library(janitor)
library(modeltime)
library(timetk)
# se here https://wilkelab.org/ggtext/
library(ggtext)
library(cowplot)
theme_set(theme_minimal())
update_geom_defaults("rect", list(fill = "midnightblue", alpha = 0.8))
update_geom_defaults("line", list(color = "midnightblue", alpha = 0.8))
#renv::snapshot()
```


```{r my theme, echo=TRUE}
theme_ben <- function(base_size = 14) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      # L'ensemble de la figure
      plot.title = element_text(size = rel(1), face = "bold", margin = margin(0,0,5,0), hjust = 0),
      # Zone où se situe le graphique
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Les axes
      axis.title = element_text(size = rel(0.85), face = "bold"),
      axis.text = element_text(size = rel(0.70), face = "bold"),
      axis.line = element_line(color = "black", arrow = arrow(length = unit(0.3, "lines"), type = "closed")),
      # La légende
      legend.title = element_text(size = rel(0.85), face = "bold"),
      legend.text = element_text(size = rel(0.70), face = "bold"),
      legend.key = element_rect(fill = "transparent", colour = NA),
      legend.key.size = unit(1.5, "lines"),
      legend.background = element_rect(fill = "transparent", colour = NA),
      # Les étiquettes dans le cas d'un facetting
      strip.background = element_rect(fill = "#17252D", color = "#17252D"),
      strip.text = element_text(size = rel(0.85), face = "bold", color = "white", margin = margin(5,0,5,0))
    )
}
```


# Problem

We have the last year seen a spike and Natural Gas prices for EU (TTF) and with a enormous volatility. 
It has left several industrices that use gas for heating and manufacturing gone out of business because they
cannot pay their bills and it has the same effect on the indicidual households.  

For many decision maker it is important to know why this is happening so they can make 
better decision on who they coorporate with. So for this analysis I would like to 
look into the reason why we are seeing this increase so a Politican can come up 
we ideas to where we can change our energy ressources. 

# Data

The data is collected from professionel in the trader field so it is in mu opinon relaiable. We can 
extract the dta from multiple places but with a fee. I have focus on the free sources which can
be bad so it is not original. Therefore the data is not comprehensive beacuse you can get better
institution where the data given can be questions.


# Process

I am going to use `R` for the cleaning.
Beacuse it is not so far a very complicated data I have only transformed the date column
to a `date class`. While importing the data it started out being a `character`.

Data is from a respect side for making analysis of the trzder market. The could be a problem
with the intrigity of the data, but if that was the case the whole side would be seen as
unreilable. Therefore it would be a bad business if they did not secure the data and check if the 
data is correct.

Because of the volatility of the Gas prices I will make a log transformation of the data
to remove some of the variance. It is a better way of removeing some of the variance so it is 
easier to remove some outlier for a time series data.

I include two new features; `year` and `month` that help us investigate if the diffrece in the prices can
be seen in the some period. 

# Analyze

* A summary of your analysis

```{r load data}
ttf <- read_csv('data/Dutch TTF Natural Gas Futures Historical Data.csv') %>% 
  janitor::clean_names()  %>% 
  select(date, price)  %>% 
  mutate(date = as.Date(date, "%m/%d/%Y") )

ttf_by_date <- ttf %>%  
  summarise_by_time(
    .date_var = date,
    .by = 'day',
    value = sum(price)
  ) %>% 
  mutate(year = as_factor(lubridate::year(date))) 
```

The Gas prices has over the last year seen a rising trend and there is several
events that has contributed to this as in seen the graph:

```{r fig.height=6}
arrows <- 
  tibble(
    x1 = as.Date(c("2022-04-01")),
    x2 = as.Date(c("2022-06-01")),
    y1 = c(300), 
    y2 = c(290)
  )

ttf_by_date %>% 
  filter(date >= '2021-01-01') %>% 
  ggplot() +
  geom_line(aes(x = date, y = value), 
            colour = '#66a182', 
            size = 0.8) +
  scale_x_date(date_labels = "%b-%y", date_breaks = '4 month') +
  labs(
    title = "<span style='font-size:15pt'><b>TTF price developemt </b> <br />
    <span style='font-size:11pt'>Historical explanation on the trend and event</span>",
    x = "", 
    y = "Euro/MWh",
    caption = "<b>Source</b>: *investment.com*"
  ) +
  theme(
    plot.title = element_markdown(lineheight = 1.1),
    legend.text = element_markdown(size = 11),
    plot.caption = element_markdown(size = 11),
    text = element_text(family="Comic Sans MS"),
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  annotate(geom = "rect", 
           xmin = as.Date("2022-06-13"), 
           xmax = as.Date("2022-08-27"),
           ymin = 0,
           ymax = 368,
           fill = "red",
           alpha = 0.1) +
  annotate(geom = "richtext",
           x = as.Date("2022-02-10"),
           y = 310, 
           size = 3,
           label = "<span style='font-family:Times New Roman'><b>Nord Stream</b> supply cuts and<br /> 
           Structural Uncertainty <span>", 
           fill = NA,
           label.color = NA,
           hjust = 0) +
  geom_curve(
    data = arrows, aes(x = x1, y = y1, xend = x2, yend = y2),
    arrow = arrow(length = unit(0.08, "inch")), size = 0.5,
    color = "gray20", curvature = 0.3) +
  annotate(geom = "richtext",
           x = as.Date("2021-12-10"),
           y = 250, 
           size = 3,
           label = 
             "<p><span style> <span style> <b>Russia</b> threatens <br/>Nord Stream supply cuts <span> </p>", 
           fill = NA,
           label.color = NA,
           hjust = 0) +
  annotate(
    geom = "segment",
    x = as.Date('2022-03-07'), 
    y = 227,
    xend = as.Date('2022-02-10'), 
    yend = 237
  ) +
  annotate(geom = "richtext",
           x = as.Date("2021-12-20"),
           y = 210, 
           size = 3,
           label = 
             "<p><span style> <span style> <b>Russia</b> <br/> invades Ukraine  <span> </p>", 
           fill = NA,
           label.color = NA,
           hjust = 0) +
  annotate(
    geom = "segment",
    x = as.Date('2022-02-24'), 
    y = 134,
    xend = as.Date('2022-01-10'), 
    yend = 199
  ) +
  annotate(geom = "richtext",
           x = as.Date("2021-08-15"),
           y = 210, 
           size = 3,
           label = 
             "<p><span style> <span style> <b>YAMAL</b> is reversed  <span> </p>", 
           fill = NA,
           label.color = NA,
           hjust = 0) +
  annotate(
    geom = "segment",
    x = as.Date('2021-12-21'), 
    y = 180,
    xend = as.Date("2021-11-20"), 
    yend = 210
  )
```


* **YAMAL Event**: The Yamal–Europe natural gas pipeline is a pipeline connecting Russian natural gas fields in the Yamal Peninsula and Western Siberia with Poland and Germany, through Belarus. In this period since Dec. 21, the link between Poland and Germany had been operating in reverse mode, carrying gas from west to east and putting upward pressure on European gas prices. 
* **Russia invades Ukraine**: TTF  price made huge gains in early trade Feb. 24 after Ukraine said Russia had launched a full-scale invasion of the country. The raise was because of the markets concern over potential disruption to Russian gas deliveries to Europe amid low storage stocks.
* **Russias thread**: Russia threatened to cut natural gas supplies to Europe via the Nord Stream 1 pipeline as part of its response to sanctions imposed over the invasion of Ukraine, a move that could heighten the turmoil in energy markets and drive consumer prices even higher.
* **Nord Stream cuts**: Russia  annoce that Nord Stream 1 would stay shut indefinitely, stoking renewed fears about shortages and gas rationing in the European Union this winter. 

After this event we have seen a decline in the prices. This is for further investigation.

# Conclusion

EU has over the years relied heavely on Russia delivering gas. After the invading and conflict EU has had no alternativ to Russia.
Therefore it is of most importance to find other ressources to either get the Gas or alternatives energy ressources.

# Appendix

## Using R enviroment `renv`

Packages alway changes so it is important to make sure that the package
you used in a start is lock on some version. 

Another benefit is for other can easy use the same packge you just have used.

For this we use the package `renv`. The first we nned to use initalise,
where we see what pakcge we use and looks into the files with code where
we have used library. It creates the dependency used for a given projects.

```
renv::init()
```

Take a and lock the package version with this function. 

```
renv::snapshot()
```

We can restore to the old packages version with `renv::``

With `.libPaths` we can see that for the projects we are using the packages in the
enviroment of renv. If we are outside of the enviroment then we wil take from the default.

See [renv documentation](https://rstudio.github.io/renv/articles/renv.html)


## Data

For Natural_Gas_Futures_Contract_1_Daily.csv bruges siden:

* https://www.eia.gov/opendata/v1/qb.php?category=462457&sdid=NG.RNGC1.D 

For storage

* [Stoage](https://agsi.gie.eu/data-overview/EU)

For Dutch TTF Natural Gas Futures Historical Data
* [TTF](https://www.investing.com/commodities/dutch-ttf-gas-c1-futures-historical-data)

For Henry Hub Gas future historia
* [Henry Hub](https://markets.businessinsider.com/commodities/natural-gas-price)


## Visual data

### By year

```{r}
ttf_by_date %>% 
  group_by(year)  %>% 
  ggplot() +
  geom_line(aes(x = date, y = value), 
            colour = '#66a182', 
            size = 0.8) +
  facet_wrap(~year, scales = "free") +
  scale_x_date(date_labels = "%b") +
  labs(
    title = "TTF price developemt by *year*  
    <span style='font-size:11pt'>Sepal width vs. sepal length for three *Iris*
    species</span>",
    x = "", 
    y = "",
    caption = "<b>Source</b>: *investment.com*"
  ) +
  theme(
    plot.title = element_markdown(lineheight = 1.1),
    legend.text = element_markdown(size = 11),
    plot.caption = element_markdown(size = 11),
    strip.text = element_textbox(
      size = 8,
      color = "white", 
      fill = "#5D729D",
      box.color = "#4A618C",
      halign = 0.5, 
      linetype = 1,
      r = unit(5, "pt"),
      width = unit(1, "npc"),
      padding = margin(2, 0, 1, 0),
      margin = margin(4, 4, 4, 4)
    )
  )
```



# Energi Market

## Spot market

The spot market is where financial instruments, such as commodities, currencies, and securities, are traded for immediate delivery. Delivery is the exchange of cash for the financial instrument

## Future market

A futures contract is a legal agreement to buy or sell a particular commodity asset, or security at a predetermined price at a specified time in the future.

## Litteratur
* [Time series book](http://home.iitj.ac.in/~parmod/document/introduction%20time%20series.pdf)
